How Large a Sample Size is Needed for Block Bootstrap Confidence Intrvals to Have Desired Coverage Rates?

### 2023-03-01 Manuscript/manuscript.tex
smoothed intro

\title{Sample Length at Which Block Bootstrapping is Effective for Estimation
  of Parameter of Time Series}

\author{Mathew Chandy\\
%   \href{mailto:mathew.chandy@uconn.edu}
% {\nolinkurl{mathew.chandy@uconn.edu}}\\
  Jun Yan\\[1ex]
  Department of Statistics, University of Connecticut\\
}
\date{}

@@

@@

\bigskip
\noindent\sc{Keywords}:
simulation;
\end{abstract}

\jy{Practice my writing tips on latex (e.g., keep line width under 80; use
  double blank lines to separate paragraphs; etc.):
  \url{
https://statds.github.io/stat-writing/using-the-right-tools-for-writing.html}}

@@

Block bootstrap is an important tool for constructing confidence intervals when
making inferences about dependent data. Early ideas were developed
independently by \citet{hall1985resampling}, \citet{carlstein1986use}, and 
\citet{kunsch1989jackknife}. % \citet{radovanov2014comparison}
It has been applied to a variety of different domains such 
as econometrics \citep{mackinnon2006bootstrap} and meteorology
\citep{varga2017generalised}. It is particularly useful for serially dependent
data when the serial dependence is not specified or of not primary interest.
As the sample size goes to infinity, block bootstrap confidence intervals are
expected to have coverage rates matching their nominal levels. For finite
sample sizes, however, a natural question is, how large a sample size is large
enough for block bootstrap confidence intervals to have the desired coverage
rates?

\citet{hesterberg2015teachers} notes that while percentile-based confidence
intervals from basic nonparametric bootstrap (simple resampling with no blocks
involved) are more accurate than t-intervals for larger sample sizes, they are
less accurate for smaller sample sizes. \citet{nevitt2001performance} find
that a sample size of 200-1000 is usually sufficient for interval estimation
using basic nonparametric bootstrap. \citet{goncalves2005bootstrap} found that
standard error estimates from block bootstrap small samples may be
significantly more accurate than inference from closed-form asymptotic
estimates. That study focused on the estimation of a coefficient parameter
within the context of linear regression. Still, the block bootstrap percentile
confidence intervals under-covered the parameter even for n = 1024.

While the aforementioned studies have looked into the necessary sample size
for various bootstrap methods to make inferences about various types of data,
the goal of this study is to find a threshold or range of sample sizes at
which block bootstrap is an effective method for estimating a parameter
specifically in the context of a time series. The method can be considered
effective if the rate at which confidence intervals created from samples of
the same size recover the parameter is approximately equal to the nominal
confidence level. This study will focus on estimating parameters of an AR(1)
process with two types of block bootstrap confidence intervals from the
literature \citep{diciccio1996bootstrap} \citep{rice2006mathematical}. How
these intervals are constructed as well as the general procedure for block
bootstrap parameter estimation is reviewed in Section 2. Section 3 offers an
explanation of the simulation design and results. Section 4 discusses the main
takeaways from the results, how these findings could be useful, and possible
future explorations.

\section{Review of Block Bootstrap}
\label{sec:blkbootreview}

>>>>

\title{How Large a Sample Size is Needed for Block Bootstrap Confidence Intrvals
  to Have Desired Coverage Rates?}

\author{Mathew Chandy
%   \href{mailto:mathew.chandy@uconn.edu}
% {\nolinkurl{mathew.chandy@uconn.edu}}\\
  \and Jun Yan\\[1ex]
  Department of Statistics, University of Connecticut\\
}
\date{}

@@

\bigskip
\noindent{\sc Keywords}:
dependent data; gsimulation
\end{abstract}

@@

\jy{Practice my writing tips on latex (e.g., keep line width under 80; use
  double blank lines to separate paragraphs; etc.):
  \url{
https://statds.github.io/stat-writing/using-the-right-tools-for-writing.html},
  chapter 2}

@@

Block bootstrap is a powerful tool for constructing confidence intervals when
making inferences about dependent data. Early ideas were developed
independently by \citet{hall1985resampling}, \citet{carlstein1986use}, and 
\citet{kunsch1989jackknife}.
It has since been applied in various fields such as econometrics
\citep{mackinnon2006bootstrap} and meteorology \citep{varga2017generalised}.
Block bootstrap is particularly useful for serially dependent
data when the serial dependence is not specified or not of primary interest.
The method is expected to produce confidence intervals with coverage rates
matching their nominal levels as the sample size grows to infinity. However,
when dealing with finite sample sizes, an important question is how large the
sample size needs to be for block bootstrap confidence intervals to have the
desired coverage rates.

The necessary sample size for bootstrap standard errors to provide valid
uncertainty measures in practice has not been extensively studied. For
independent data and non-block bootstrap, \citet{hesterberg2015teachers} notes
that while percentile-based confidence intervals from nonparametric bootstrap
are more accurate than $t$-intervals for larger sample sizes, they are
less accurate for smaller sample sizes. In the contect of structural equation
modeling, \citet{nevitt2001performance} find
that a sample size of 200--1000 is usually sufficient for interval estimation
using standard nonparametric bootstrap. For dependent data, the necessary sample
size for block bootstrap is even less clear. In a linear regression setting with
dependent data, \citet{goncalves2005bootstrap} find that, standard error
estimates from block bootstrap in small samples may be more accurate than
inference from closed-form asymptotic estimates, but block bootstrap percentile
confidence intervals still do not have sufficient coverage even for sample size
$n = 1024$.

The goal of this paper is to provide some practical recommendations on necessary
sample size for block bootstrap with dependent data. We consider a simple
situation of a stationary time series, where the parameters of interests are the
mean, standard deviation, and the first-order autocorrelation coefficient.
We compare four variants of confidence intervals from block bootstrap:
\jy{list them} literature \citep{diciccio1996bootstrap}
\citep{rice2006mathematical}.
Their empirical coverage rates at different sample sizes and dependence levels
are compared in a simulation study.
\jy{may reveal the most interesting result here to engage interests.}

The remainder of the papers is organized as follows.
\jy{refer to sections by label instead of hard number}
these intervals are constructed as well as the general procedure for block
bootstrap parameter estimation is reviewed in Section 2. Section 3 offers an
explanation of the simulation design and results. Section 4 discusses the main
takeaways from the results, how these findings could be useful, and possible
future explorations.

\section{Block Bootstrap Confidence Intervals}
\label{sec:blkbootreview}
